\documentclass[11pt,twoside]{NJOYMan}
\usepackage{fancyhdr}
\usepackage[letterpaper, inner=1.5in,outer=1.0in,bottom=1.25in,top=1.25in]
{geometry}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{array}
\usepackage{longtable}
\usepackage{amsmath}
\usepackage[hang,normalsize]{caption}

%\usepackage{tocloft}
\usepackage{tocbibind}
\usepackage{makeidx}
\usepackage[totoc]{idxlayout}

\setlongtables

\pagestyle{fancy}

\newcommand{\rr}{\raggedright}
\newcommand{\tn}{\tabularnewline}

\newcommand{\ty}{\tt}

% This provides more subsection indent spacing for toc titles:
\makeatletter
\renewcommand\l@subsection{\@dottedtocline{2}{1.6em}{2.8em}}
\makeatother

% The following gets the document number into the title page header,
% ... but nothing else
\fancypagestyle{plain}{
\fancyhf{}                             %clear headers and footers
\fancyhead[R]{LA-UR-17-20093}          %

\renewcommand{\headrulewidth}{0pt}     %no line across the top of this page
\renewcommand{\footrulewidth}{0pt}}    %no line across the bottom of this page

% Some stuff from the internet to get better figure placement:
\renewcommand{\topfraction}{0.95}      % max fraction of floats at top
\renewcommand{\bottomfraction}{0.95}   % max fraction of floats at bottom
%   Parameters for TEXT pages (not float pages):
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{2}            % 2 may work better
\setcounter{dbltopnumber}{2}           % for 2-column pages
\renewcommand{\dbltopfraction}{0.9}    % fit big float above 2-col. text
\renewcommand{\textfraction}{0.05}     % allow minimal text w. figs


\title{\bf The NJOY Nuclear Data Processing System,\\Version 2016}
%\author{R. E. MacFarlane and A. C. Kahler\\
%  Theoretical Division\\
%  Los Alamos National Laboratory}
\author{\\
  Original Author:  R. E. MacFarlane\\
  Theoretical Division\\
  Los Alamos National Laboratory\\
\\
\\
Contributing Authors\\
D. W. Muir\\
R. M Boicourt\\
A. C. Kahler\\
J. L. Conlin\\
W. Haeck\\
\\
\\
Current Editor:  A. C. Kahler\\
\\
\\
Original Issue:  December 19, 2016\\
\\
\\
Updated for NJOY2016.39
\\}

\date{July 3, 2018}

\makeindex

\includeonly{
    intro,
    njoyx,
    reconx,
    broadx,
    unresx,
    heatx,
    thermx,
    groupx,
    gaminx,
    errorx,
    covx,
    modex,
    dtfx,
    ccccx,
    matxsx,
    resxsx,
    acex,
    powx,
    wimsx,
    plotx,
    viewx,
    mixx,
    purx,
    leapx,
    gaspx,
    mandtx
}

\newcommand{\exx}[1]{{\rm e}^{\textstyle #1}}
\newcommand{\intall}{\int_{-\infty}^\infty}

\raggedbottom
\begin{document}

\maketitle
\abstract{The NJOY Nuclear Data Processing System, version 2016,
is a comprehensive computer code package for producing pointwise
and multigroup cross sections and related quantities from evaluated
nuclear data in the ENDF-4 through ENDF-6 legacy cardimage
formats.  NJOY works with evaluated files for incident neutrons,
photons, and charged particles, producing libraries for a wide
variety of particle transport and reactor analysis codes.}

% Only want a footer on the second page
% (that will contain the LANL/government disclaimer
% and copyright statements).
\clearpage
\fancyhf{}                           %Clear any header/footer definitions
%
\renewcommand{\headrulewidth}{0pt}   %no line across the top of this page
\renewcommand{\footrulewidth}{0pt}   %no line across the bottom of this page

%
\begin{center} \textbf{Disclaimer of Liability:} \end{center}
Neither the United States Government nor the Los Alamos
National Security, LLC., nor any of their employees, makes any warranty,
express or implied, including the warranties of merchantability and
fitness for a particular purpose, or assumes any legal liability or
responsibility for the accuracy, completeness, or usefulness of any
information, apparatus, product, or process disclosed, or represents
that its use would not infringe privately owned rights.

\vspace{0.25in}
\begin{center} \textbf{Disclaimer of Endorsement:} \end{center}
Reference herein to any specific commercial products,
process, or service by trade name, trademark, manufacturer, or otherwise,
does not necessarily constitute or imply its endorsement, recommendation,
or favoring by the United States Government or the Los Alamos National
Security, LLC. The views and opinions of authors expressed herein do not
necessarily state or reflect those of the United States Government or
the Los Alamos National Security, LLC., and shall not be used for
advertising or product endorsement purposes.

\vspace{0.25in}
\begin{center} \textbf{Copyright Notice:} \end{center}
Copyright 2016. Los Alamos National Security, LLC. This software was
produced under U.S. Government contract DE-AC52-06NA25396 for
Los Alamos National Laboratory (LANL), which is operated by
Los Alamos National Security, LLC for the U.S. Department of
Energy. The U.S. Government has rights to use, reproduce, and
distribute this software.  NEITHER THE GOVERNMENT NOR LOS ALAMOS
NATIONAL SECURITY, LLC MAKES ANY WARRANTY, EXPRESS OR IMPLIED, OR
ASSUMES ANY LIABILITY FOR THE USE OF THIS SOFTWARE.  If software
is modified to produce derivative works, such modified software
should be clearly marked, so as not to confuse it with the version
available from LANL.

Additionally, redistribution and use in source and binary forms, with
or without modification, are permitted provided that the following
conditions are met:
\begin{enumerate}
\item Redistributions of source code must retain the above copyright
notice, this list of conditions and the following disclaimer.
\item Redistributions in binary form must reproduce the above
copyright notice, this list of conditions and the following disclaimer
in the documentation and/or other materials provided with the
distribution.
\item Neither the name of Los Alamos National Security, LLC, Los
Alamos National Laboratory, LANL, the U.S. Government, nor the
names of its contributors may be used to endorse or promote products
derived from this software without specific prior written permission.
\end{enumerate}
THIS SOFTWARE IS PROVIDED BY LOS ALAMOS NATIONAL SECURITY,
LLC AND CONTRIBUTORS "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES,
INCLUDING, BUT NOT LIMITED TO, THE IMPLIED WARRANTIES OF
MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE
DISCLAIMED. IN NO EVENT SHALL LOS ALAMOS NATIONAL SECURITY,
LLC OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF
USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED
AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT
LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN
ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
POSSIBILITY OF SUCH DAMAGE.

\vspace{7.1in}
\begin{center}
\copyright Copyright 2016 Los Alamos National Security,
LLC All Rights Reserved
\end{center}
%
% End of disclaimer page stuff.

\newpage

% Define left and right page headers and footers:
\fancyhf{}
\fancyhead[RO,LE]{LA-UR-17-20093}
\fancyhead[RE,LO]{\slshape \leftmark}
\fancyfoot[RO,LE]{\thepage}
%
\fancyfoot[LO,RE]{NJOY2016}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0pt}

% Use lower case roman numerals for table of contents, list of figures
% and list of table page numbers.
% Start with page iv since we already have title and two disclaimer pages.
\pagenumbering{roman}
\setcounter{page}{4}

%\renewcommand\contentsname{}
%if active, will delete "CONTENTS" from header & toc
\tableofcontents
\newpage
\listoffigures
\newpage
\listoftables
\cleardoublepage %make sure document starts on an "odd" page.

% Now switch to standard page numbers, and reset to page 1.
\pagenumbering{arabic}
\setcounter{page}{1}

% Standard line spacing command.
% Can use \begin{singlespace} ... \end{singlespace}  (or use doublespace)
% within the .tex files as needed.
\onehalfspacing  %looks nice to me for the primary setting ...
\normalsize

%
% Each include file ends with a \cleardoublepage command
% ... which inserts a blank even numbered page, if necessary
\include{intro}
\include{njoyx}
\include{reconx}
\include{broadx}
\include{unresx}
\include{heatx}
\include{thermx}
\include{groupx}
\include{gaminx}
\include{errorx}
\include{covx}
\include{modex}
\include{dtfx}
\include{ccccx}
\include{matxsx}
\include{resxsx}
\include{acex}
\include{powx}
\include{wimsx}
\include{plotx}
\include{viewx}
\include{mixx}
\include{purx}
\include{leapx}
\include{gaspx}
\include{mandtx}

%\section{References}
%comment out so "References" doesn't get a section number in the TOC
\begin{thebibliography}{99}
\begin{singlespace}
\bibitem{NJ1} R. E. MacFarlane and R. M. Boicourt, ``NJOY:  A
  Neutron and Photon Processing System,'' {\it Trans. Am. Nucl.
  Soc.} \underline{22}, 720 (1975).

\bibitem{NJ2} R. E. MacFarlane, D. W. Muir, and R. M. Boicourt,
  ``The NJOY Nuclear Data Processing System, Volume 1:
  User's Manual,'' Los Alamos National Laboratory report
  LA-9303-M, Vol. I (ENDF-324), (May 1982).  R. E. MacFarlane,
  D. W. Muir, and R. M. Boicourt, ``The NJOY Nuclear Data
  Processing System, Volume II: The NJOY, RECONR, BROADR,
  HEATR, and THERMR Modules,'' Los Alamos National Laboratory
  report LA-9303-M, Vol. II (ENDF-324) (May 1982).
  R. E. MacFarlane and D. W. Muir, ``The NJOY Nuclear Data
  Processing System, Volume III: The GROUPR, GAMINR, and
  MODER Modules,'' Los Alamos National Laboratory report
  LA-9303-M, Vol. III (ENDF-324) (October 1987).  D. W. Muir
  and R. E. MacFarlane, ``The NJOY Nuclear Data Processing
  System, Volume IV: The ERRORR and COVR Modules,'' Los Alamos
  National Laboratory report LA-9303-M, Vol. IV (ENDF-324)
  (December 1985).

\bibitem{NJOY91} R. E. MacFarlane and D. W. Muir, ``The NJOY
  Nuclear Data Processing System, Version 91,'' Los Alamos National
  Laboratory report LA-12740-M (1994).

\bibitem{NJ10} R. E. MacFarlane and A. C. Kahler, ``Methods
 for Processing ENDF/B-VII with NJOY," {\it Nuclear Data Sheets},
 {\bf 111}, 2739--2890 (2010).

\bibitem{NJ12} A. C. Kahler, editor, ``The NJOY Nuclear Data
  Processing System, Version 2012,'' Los Alamos National
  Laboratory report LA-UR-12-27079 (2012).

\bibitem{njoytutorial} For an online tutorial on NJOY, see
 \href{http://t2.lanl.gov}{http://t2.lanl.gov}, and follow links
 to the ``Nuclear Information Service (NIS)", ``Training area"
 and ``NJOY" which leads to ``Understanding NJOY,'' Los Alamos
 National Laboratory report LA-UR-00-1538 (March 2000).

\bibitem{ENDF7} M. B. Chadwick {\it et al.}, ``ENDF/B-VII.0:
  Next Generation Evaluated Nuclear Data Library for Nuclear
  Science and Technology,'' {\it Nuclear Data Sheets}, {\bf 107},
  2931--3060 (2006).

\bibitem{ENDF71} M. B. Chadwick {\it et al.}, ``ENDF/B-VII.1 Nuclear
  Data for Science and Technology:  Cross Sections, Covariances,
  Fission Product Yields and Decay Data,'' {\it Nuclear Data Sheets},
  {\bf 112}, 2887--2996 (2011).

\bibitem{ENDF102} A. Trkov, M. Herman and D. A. Brown, Eds., “ENDF-6
  Formats Manual, Data Formats and Procedures for the Evaluated Nuclear
  Data Files ENDF/B-VI and ENDF/B-VII,” Brookhaven National Laboratory
  report BNL-90365-2009 Rev.2 (CSEWG Document
  ENDF-102) (October 24, 2012). For the latest version of
  this report, see \href{http://www.nndc.bnl.gov}{http://www.nndc.bnl.gov}
  and take the “CSEWG” tab to “ENDF-6 Formats Manual.”

\bibitem{endftutorial} For an online tutorial on the ENDF formats,
 see \href{http://t2.lanl.gov}{http://t2.lanl.gov}, and follow links
 to the ``Nuclear Information Service (NIS)", ``Training area"
 and ``ENDF" which leads to ``An Introduction to the ENDF Formats,"
 Los Alamos National Laboratory report LA-UR-98-1779 (January 1998).

\bibitem{CCCC4} R. D. O'Dell, ``Standard Interface Files and
  Procedures for Reactor Physics Codes, Version IV,'' Los Alamos
  Scientific Laboratory report LA-6941-MS (September 1977).

\bibitem{TRANSX} R. E. MacFarlane,  ``TRANSX-CTR: A Code for
  Interfacing MATXS Cross-Section Libraries to Nuclear Transport
  Codes for Fusion Systems Analysis,'' Los Alamos National
  Laboratory report LA-9863-MS (February 1984).

\bibitem{ANISN} W. W. Engle, Jr., ``A Users Manual for ANISN,
  A One-Dimensional Discrete Ordinates Transport Code with
  Anisotropic Scattering,'' Oak Ridge Gaseous Diffusion Plant
  Computing Technology Center report K-1693 (1967).

\bibitem{ONEDANT} R. D O'Dell, F. W. Brinkley, Jr.,
  D. R. Marr, and R. E. Alcouffe, ``Revised User's
  Manual for ONEDANT: A Code Package for One-Dimensional,
  Diffusion-Accelerated, Neutral-Particle Transport,''
  Los Alamos National Laboratory report LA-9184-M, Rev.
  (December 1989).

\bibitem{TWODANT} R. E. Alcouffe, F. W. Brinkley, D. R. Marr,
  and R. D. O'Dell, ``User's Guide for TWODANT: A Code Package
  for Two-Dimensional, Diffusion-Accelerated, Neutral-Particle
  Transport,'' Los Alamos National Laboratory report
  LA-10049 Revised (February 1990).

\bibitem{DIF3D} K. L. Derstine, ``DIF3D: A Code to Solve One-,
  Two-, and Three-Dimensional Finite-Difference Diffusion Theory
  Problems,'' Argonne National Laboratory report ANL-82-64
  (April 1984).

\bibitem{PARTISN} R. E. Alcoufe, R. S. Baker, J. A. Dahl, S. A. Turner,
  and R. C. Ward, ``PARTISN: A Time-Dependent, Parallel Neutral Particle
  Transport Code System," Los Alamos National Laboratory report
  LA-UR-05-3925 (Revised May 2005).

\bibitem{MCNP} X-5 Monte Carlo Team, ``MCNP--A General Monte Carlo
  N-Particle Transport Code, Version 5," Los Alamos National
  Laboratory report LA-UR-03-1987 (April 2003).

\bibitem{WIMS} J. R. Askew, F. J. Fayers, and P. B. Kemshell,
  ``A General Description of the Lattice Code WIMS,''
  {\it J. Brit. Nucl. Energy Soc.} \underline{5}, 564 (1966).

\bibitem{webinclude} Userinput information may be found on
 the NJOY homepage, found under the ``Nuclear Information
 Service (NIS)", ``Codes" and ``NJOY" links at
 \href{http://t2.lanl.gov}{http://t2.lanl.gov}.

\bibitem{MINX} C. R. Weisbin, P. D. Soran, R. E. MacFarlane,
  D. R. Harris, R. J. LaBauve, J. S. Hendricks, J. E. White,
  and R. B. Kidman, ``MINX, A Multigroup Interpretation of
  Nuclear X-Sections from ENDF/B,'' Los Alamos Scientific
  Laboratory report LA-6486-MS (ENDF-237) (September 1976).

\bibitem{LAPHAN0} D. J. Dudziak, R. E. Seamon, and D. V. Susco,
  ``LAPHAN0:  A Multigroup Photon-Production Matrix and Source
  Code for ENDF,'' Los Alamos Scientific Laboratory report
  LA-4750-MS (ENDF-156) (January 1972).

\bibitem{GAMLEG} K. D. Lathrop, ``GAMLEG--A FORTRAN Code to
  Produce Multigroup Cross Sections for Photon Transport Calculations,''
  Los Alamos Scientific Laboratory report LA-3267 (April 1965).

\bibitem{DTF} K. D. Lathrop, ``DTF-IV, A FORTRAN Program for
  Solving the Multigroup Transport Equation with Anisotropic
  Scattering,'' Los Alamos Scientific Laboratory report
  LA-3373 (November 1965).

\bibitem{ETOPL} R. J. LaBauve, C. R. Weisbin, R. E. Seamon,
  M. E. Battat, D. R. Harris, P. G. Young, and M. M. Klein,
  ``PENDF: A Library of Nuclear Data for Monte Carlo
  Calculations Derived from Data in the ENDF/B Format,''
  Los Alamos Scientific Laboratory report LA-5687 (October 1974).

\bibitem{RESEND} O. Ozer, Ed. ``Description of the ENDF/B
  Processing Codes and Retrieval Subroutines,'' Brookhaven
  National Laboratory report BNL-50300 (ENDF-11) (June 1971).

\bibitem{SIGMA1} D. E. Cullen, ``Program SIGMA1 (Version 74-1),''
  Lawrence Livermore Laboratory report UCID-16426 (January 1974).

\bibitem{ETOX} R. E. Schenter, J. L. Baker, and R. B. Kidman,
  ``ETOX, A Code to Calculate Group Constants for Nuclear Reactor
  Calculations,'' Battelle Northwest Laboratory report
  BNWL-1002 (1969).

\bibitem{CCCC3} B. M. Carmichael, ``Standard Interface Files and
  Procedures for Reactor Physics Codes, Version III,'' Los Alamos
  Scientific Laboratory report LA-5486-MS (1974).

\bibitem{MACK} M. A. Abdou, C. W. Maynard, and R. Q. Wright,
  ``MACK: A Computer Program to Calculate Neutron Energy Release
  Parameters (Fluence to Kerma Factors) and Multigroup
  Neutron Reaction Cross Sections from Nuclear Data in ENDF
  Format,'' Oak Ridge National Laboratory report ORNL-TM-3994
  (1973).

\bibitem{FLANGE} H. C. Honek and D. R. Finch, ``FLANGE-II
  (Version 71-1), A Code to Process Thermal Neutron Data frp, an
  ENDF/B Tape,'' E. I. DuPont de Nemours and Co. Savannah River
  Laboratory report DP-1278 (1971).

\bibitem{HEXSCAT} Y. D. Naliboff and J. U. Koppel, ``HEXSCAT,
  Coherent Elastic Scattering of Neutrons by Hexagonal Lattices,''
  General Atomic report GA-6026 (1964).

\bibitem{NJ78} R. E. MacFarlane, R. J. Barrett, D. W. Muir, and
  R. M. Boicourt, ``NJOY: A Comprehensive ENDF/B Processing System,''
  in A Review of Multigroup Nuclear Cross-Section Processing,
  Proceedings of a Seminar-Workshop, Oak Ridge, Tennessee,
  March 14-16, 1978,'' Oak Ridge National Laboratory report
  ORNL/RSIC-41 (October 1978).

\bibitem{NJ89} R. E. MacFarlane, ``Introducing NJOY89,'' in
  Proceedings of the Seminar on NJOY and Themis, 20-21 June 1989,
  OECD/NEA Data Bank, Saclay, France (1989).  See also ``Nuclear
  Theory and Application Progress Report, January 1, 1989---April
  1, 1990,'' pp. 21-38, Los Alamos National Laboratory report
  LA-11972-PR (December 1990).

\bibitem{ERRORJ} Go Chiba, ``ERRORJ, A Code to Process Neutron-Nuclide
  Reaction Covariance, Version 2.3,'' JAEA-Data/Code 2007-007, March 2007.

\bibitem{SAMMY} N.M.Larson, ``Updated Users’ Guide for SAMMY:
  Multilevel R-Matrix Fits to Neutron Data Using Bayes’ Equations,''
  Oak Ridge National Laboratory report ORNL/TM-9179/R8 (October 2008).

\bibitem{MC2} B. J. Toppel, A. L. Rago, and D. M. O'Shea,
  ``MC$^2$, A Code to Calculate Multigroup Cross Sections,''
  Argonne National Laboratory report ANL-7318 (1967).

\bibitem{MC22} H. Henryson II, B. J. Toppel, and C. G. Stenberg,
  ``MC$^2$-2: A Code to Calculate Fast Neutron Spectra and Multigroup
  Cross Sections,'' Argonne National Laboratory report ANL-8144
  (ENDF-239) (1976).

\bibitem{Bondarenko} I. I. Bondarenko, Ed., {\it Group Constants for Nuclear
  Reactor Calculations} (Consultants Bureau, New York, 1964).

\bibitem{Kidman} R. B. Kidman, ``Cross Section Structure Factor
   Interpolation Schemes,'' Hanford Engineering Development
   Laboratory report HEDL-TME-71-40 (1971).

\bibitem{TRANSX2} R. E. MacFarlane, ``TRANSX 2: A Code for Interfacing
   MATXS Cross-Section Libraries to Nuclear Transport Codes,''
   Los Alamos National Laboratory report LA-12312-MS (July 1992).

\bibitem{Muir} D. W. Muir, ``Gamma Rays, Q-Values, and Kerma Factors,''
   Los Alamos Scientific Laboratory report LA-6258-MS (March 1976).

\bibitem{ebal} R. E. MacFarlane, ``Energy Balance of ENDF/B-V,''
   Trans. Am. Nucl. Soc. \underbar{33}, 681 (1979).  See also
   R. E. MacFarlane, ``Energy Balance of ENDF/B-V.2,''
   Minutes of the Cross Section Evaluation Working Group
   (available from the National Nuclear Data Center, Brookhaven
   National Laboratory, Upton, NY) (May 1984).

\bibitem{Gabriel} T. A. Gabriel, J. D. Amburgy, and N. M. Greene,
   ``Radiation-Damage Calculations: Primary Knock-On Atom Spectra,
   Displacement Rates, and Gas Production Rates,''
   {\it Nucl. Sci. Eng.} \underline{61}, 21 (1976).

\bibitem{Doran} D. G. Doran, ``Neutron Displacement Cross Sections for
   Stainless Steel and Tantalum Based on a Linhard Model,'' {\it Nucl. Sci.
   Eng.} \underline{49}, 130 (1972).

\bibitem{Greenwood} L. R. Greenwood and R. K. Smither, ``Displacement
   Damage Calculations with ENDF/B-V,'' in Proceedings of the Advisory Group
   Meeting on Nuclear Data for Radiation Damage Assessment and Reactor
   Safety Aspects, October 12-16, 1981, IAEA, Vienna, Austria (October 1981).

\bibitem{Robinson} M. T. Robinson, in {\it Nuclear Fusion Reactors}
   (British Nuclear Energy Society, London, 1970).

\bibitem{Lindhard} J. Lindhard, V. Nielsen, M. Scharff, and P. V. Thomsen,
   {\it Kgl. Dansk, Vidensk. Selsk, Mat-Fys. Medd.} \underline{33} (1963).

\bibitem{alb} R. E. MacFarlane, D. W. Muir, and F. M. Mann,
   ``Radiation Damage Calculations with NJOY,''
   {\it J. Nucl. Materials} \underline{122} and \underline{123},
   1041 (1984).

\bibitem{advanced} R. E. MacFarlane and D. G. Foster, Jr.,
   ``Advanced Nuclear Data for Radiation Damage Calculations,''
   {\it J. Nucl. Materials} \underline{122} and \underline{123},
   1047 (1984).

\bibitem{old102} R. Kinsey, Ed., ``ENDF-102, Data Formats and Procedures
   for the Evaluated Nuclear Data File, ENDF,'' Brookhaven National
   Laboratory report BNL-NCS-50496 (ENDF-102) 2nd Edition
   (ENDF/B-V) (October 1979).

\bibitem{GAreport} J. U. Koppel and D. H. Houston, ``Reference Manual
   for ENDF Thermal Neutron Scattering Data,'' General Atomic report
   GA-8774 revised and reissued as ENDF-269 by the National Nuclear Data
   Center, Brookhaven National Laboratory (1978).

\bibitem{ref5} G. I. Bell and S. Glasstone, {\it Nuclear Reactor Theory}
  (Van Nostrand Reinhold, New York, 1970).

\bibitem{1DX} R. W. Hardie and W. W. Little, Jr., ``1DX, A One-Dimensional
  Diffusion Code for Generating Effective Nuclear Cross Sections,''
  Battelle Northwest Laboratory report BNWL-954 (1969).

\bibitem{ref9} R. E. MacFarlane, ``ENDF/B-IV and -V Cross Sections for Thermal
  Power Reactor Analysis,'' in Proc. Intl. Conf. of Nuclear Cross Sections
  for Technology, Knoxville, TN (October 22-26 1979), National Bureau of
  Standards Publication 594 (September 1980).

\bibitem{ref10} G. I. Bell, G. E. Hansen, and H. A. Sandmeier, ``Multitable
  Treatments of Anisotropic Scattering in S$_{\rm N}$ Multigroup
  Transport Calculations,'' Nucl. Sci. Eng. \underbar{28}, 376 (1967).

\bibitem{ref13}  M. Abramowitz and I. Stegun, {\it Handbook of Mathematical
  Functions} (Dover Publications, New York, 1965).

\bibitem{ref14}  D. G. Madland, ``New Fission Neutron Spectrum
  Representation for ENDF,'' Los Alamos National Laboratory
  report LA-9285-MS (ENDF-321) (April 1982).

\bibitem{km} C. Kalbach and F. M. Mann, ``Phenomenology of Continuum
   Angular Distributions. I. Systematics and Parameterization,''
   Phys. Rev. C. \underline{23}, 112 (1981).

\bibitem{k86} C. Kalbach, ``Systematics of Continuum Angular Distributions:
  Extensions to Higher Energies,''  {\it Phys. Rev. C.}
  \underline{37}, 2350 (1988).

\bibitem{GNASH} P. G. Young and E. D. Arthur, ``GNASH: A Preequilibrium
  Statistical Nuclear Model Code for Calculation of Cross Sections and
  Emission Spectra,''  Los Alamos Scientific Laboratory report LA-6947 (1977).

\bibitem{powr} R. E. MacFarlane, ``ENDF/B-V Cross-Section
  Library for Reactor Cell Analysis,'' Electric Power Research
  Institute report EPRI NP-3418 (February 1984).

\bibitem{LIBIV} R. B. Kidman and R. E. MacFarlane, ``LIB-IV, A Library
  of Group Constants for Nuclear Reactor Calculations,'' Los Alamos
  Scientific Laboratory report LA-6260-MS (1976).

\bibitem{one} G. Lehner, ``Reaction Rates and Energy Spectra
  for Nuclear Reactions in High Energy Plasmas,'' {\it Z. Physik}
  {\bf 232}, 174 (1970).

\bibitem{two} D. E. Nagle, W. E. Quinn, W. B. Riesenfeld,
  and W. Leland, ``Ion Temperature in Scylla, as Determined from the
  Reaction D(d,p)T,'' {\it Phys. Rev. Letters} {\bf 3}, 318 (1959).

\bibitem{three} N. Jarmie and J. D. Seagrave (Eds.),
  ``Charged Particle Cross Sections,'' Los Alamos Scientific
  Laboratory report LA-2014 (1956).

\bibitem{four} D. W. Muir, ``Sensitivity of Neutron Multigroup
  Cross Sections to Thermal Broadening of the Fusion Peak,''
  Proc. 1st Topical Meeting on the Technology of Controlled
  Nuclear Fusion, San Diego, Calif., April 1974,
  CONF-740402-P2, p.166 (1975).

\bibitem{HUGO}  R. W. Roussin, J. R. Knight, J. H. Hubbell, and R. J. Howerton,
  ``Description of the DLC-99/HUGO Package of Photon Interaction Data in
  ENDF/B-V Format,'' Oak Ridge National Laboratory report ORNL/RSIC-46
  (ENDF-335) (December 1983).

\bibitem{Hubbell} J. H. Hubbell, Wm. J. Viegle, E. A. Briggs,
  R. T. Brown, D. T. Cromer, and R. J. Howerton, ``Atomic Form Factors,
  Incoherent Scattering Functions, and Photon Scattering Cross Sections,''
  {\it J. Phys. Chem. Ref. Data} {\bf 4}, 471 (1975).

\bibitem{Gerstl} S. A. W. Gerstl, D. J. Dudziak, and D. W. Muir,
  ``Cross Section Sensitivity and Uncertainty Analysis with Application
  to a Fusion Reactor,'' {\it Nucl. Sci. Eng.} {\bf 62}, 137 (January 1977).

\bibitem{Reupke} W. A. Reupke, D. W. Muir, and J. N. Davidson,
  ``Consistency of Neutron Cross-Section Data, S$_{\rm{n}}$ Calculations,
  and Measured Tritium Production for a 14-MeV Neutron-Driven Sphere of
  Natural-Lithium Deuteride,'' {\it Nucl. Sci. Eng.} {\bf 82}, 416 (December
  1982).

\bibitem{Hamilton} D. W. Muir, ``Evaluation of Correlated Data Using
  Partitioned Least Squares: A Minimum-Variance Derivation,'' {\it Nucl.
  Sci. and Eng.} {\bf 101}, 88-93 (January 1989).

\bibitem{Weisbin} C. R. Weisbin, E. M. Oblow, J. Ching, J. E. White, R.
  Q. Wright, and J. Drischler, ``Cross Section and Method Uncertainties:
  The Application of Sensitivity Analysis to Study Their Relationship in
  Radiation Transport Benchmark Problems,'' Oak Ridge National Laboratory
  report ORNL-TM-4847 (ENDF-218) (August 1975).  (See especially Chapter IV.)

\bibitem{Smith} J. D. Smith III, ``Processing ENDF/B-V Uncertainty Data
  into Multigroup Covariance Matrices,'' Oak Ridge National Laboratory
  report ORNL/TM-7221 (ENDF-295) (June 1980).

\bibitem{Young-2} P. G. Young, ``Evaluation of n + $^7$Li Reactions
  using Variance-Covariance Techniques,'' {\it Trans. Am. Nucl. Soc.}
  {\bf 39}, 272 (1981).

\bibitem{Young} P. G. Young, J. W. Davidson, and D. W. Muir,
  ``Evaluation of the $^7$Li(n,n't)$^4$He Cross Section for ENDF/B-VI and
  Application to Uncertainty Analysis,'' {\it Fusion Tech.} {\bf 15},
  440-448 (1989).

\bibitem{LaBauve} D. W. Muir and R. J. LaBauve, ``COVFILS: A 30-Group
  Covariance Library Based on ENDF/B-V,'' Los Alamos National Laboratory
  report LA-8733-MS (ENDF-306)  (March 1981).

\bibitem{COVFILS2} D. W. Muir, ``COVFILS-2: Neutron Data and Covariances
  for Sensitivity and Uncertainty Analysis,'' {\it Fusion Tech.} {\bf 10}
  (3), Part 2B, 1461 (November 1986).

\bibitem{sensit} S. A. W. Gerstl, ``SENSIT: A Cross-Section and Design
  Sensitivity and Uncertainty Analysis Code,'' Los Alamos Scientific
  Laboratory report LA-8498-MS (August 1980).

\bibitem{sensit-2} M. J. Embrechts, ``SENSIT-2D: A Two-Dimensional
  Cross-Section Sensitivity and Uncertainty Analysis Code,'' Los Alamos
  National Laboratory report LA-9515-MS (October 1982).

\bibitem{multigroup} D. W. Muir, R. E. MacFarlane, and R. M. Boicourt,
  ``Multigroup Processing of ENDF/B Dosimetry Covariances,'' Proc. 4th
  ASTM-EURATOM Symp. on Reactor Dosimetry, Gaithersburg, Maryland, March
  22-26, 1982, NUREG/CP-0029 (CONF-820321), p. 655 (1982).

\bibitem{CLAW4} R. J. Barrett and R. E. MacFarlane, ``Coupled
  Neutron and Photon Cross Sections for Transport Calculations,''
  Los Alamos Scientific Laboratory report LA-7808-MS (April 1979).

\bibitem{LINXBINX} R. E. MacFarlane and R. B. Kidman,
  ``LINX and BINX: CCCC Utility Codes for the MINX Multigroup
  Processing Code,'' Los Alamos Scientific Laboratory report
  LA-6219-MS (February 1976).

\bibitem{CINX} ``CINX: Collapsed Interpretation of Nuclear
  X Sections,'' Los Alamos Scientific Laboratory
  report LA-6287-MS (April 1976).

\bibitem{SPHINX} W. J. Davis, M. B. Yarbrough, and A. B. Bortz,
  ``SPHINX: A One Dimensional Diffusion and Transport Nuclear
  Cross Section Processing Code,'' Westinghouse Advanced Reactors
  Division report WARD-XS-3045-17 (August 1977).

\bibitem{TDOWN} C. L. Cowan, B. A. Hutchins, and J. E. Turner,
  ``TDOWN -- A Code to Generate Composition and Spatially
  Dependent Cross Sections,'' General Electric report GEAP-13740
  (August 1971).

\bibitem{DANDE} R. J. LaBauve, T. R. England, D. C. George,
  R. E. MacFarlane, and W. B. Wilson, ``DANDE---A Linked Code
  System for Core Neutronics/Depletion Analysis,'' Los Alamos
  National Laboratory report LA-10412-MS (June 1985).

\bibitem{UPD} R. E. MacFarlane and D. C. George, ``UPD: A
  Portable Version-Control Program,'' Los Alamos National Laboratory
  report LA-12057-MS (April 1991).

\bibitem{KAOS5} Y. Farawila, Y. Gohar, and C. Maynard, ``KAOS/LIB-V:
  A Library of Nuclear Response Functions Generated by KAOS-V Code
  From ENDF/B-V and Other Data Files" Argonne National Laboratory
  report ANL/FPP/TM-241 (April 1989).

\bibitem{MCN} E. D. Cashwell, J. R. Neergaard, W. M. Taylor, and
  G. D. Turner, ``MCN: A Neutron Monte Carlo Code," Los Alamos Scientific
  Laboratory report LA-4751 (January 1972).

\bibitem{MCG} E. D. Cashwell, J. R. Neergaard, C. J. Everett,
  R. G. Schrandt, W. M. Taylor, and G. D. Turner, ``Monte Carlo Photon
  Codes MCG and MCP," Los Alamos Scientific Laboratory report
  LA-5157-MS (March 1973).

\bibitem{MCP} C. J. Everett and E. D. Cashwell, ``MCP Code
  Fluorescence-Routine Revision," Los Alamos Scientific Laboratory
  report LA-5240-MS (May 1973).

\bibitem{redbook} Adobe Systems Incorporated, {\it Postscript
  Language Reference Manual}, Addison-Wesley Publishing Company, 1985.

\bibitem{levitt} L. B. Levitt, ``The Probability Table Method
  for Treating Unresolved Resonances in Monte Carlo Criticality
  Calculations,'' {\it Trans. Am. Nucl. Soc.} \underline{14}, 648 (1971).

\bibitem{vim} R. E. Prael and L. J. Milton, ``A User's Manual
  for the Monte Carlo Code VIM,'' Argonne National Laboratory
  report FRA-TM-84 (1976).  Also, R. Blomquist, ``VIM continuous
  energy Monte Carlo transport code,'' in Proc. Inter. Conf. on
  Mathematics, Computations, Reactor Physics and Environmental
  Analysis, Portland, OR, April 30 -- May 4, 1995.

\bibitem{GASKET} J. U. Koppel, J. R. Triplett, and Y. D. Naliboff,
  ``GASKET: A Unified Code for Thermal Neutron Scattering,''
  General Atomics report GA-7417 (Rev.) (March 1967).

\bibitem{McLatchie} R. C. F. McLatchie, 1962, unpublished.

\bibitem{Butland} A. T. Butland, ``LEAP and ADDELT, A Users Guide
  to Two Complementary Codes on the ICL-470 for Calculating The Scattering
  Law From a Phonon Frequency Function,''       Atomic Energy Establishment
  Winfrith report AEEW-M-1200, 1973.

\bibitem{Picton} D. J. Picton, personal communication to G. J. Russell, 1987.

\bibitem{New} R. E. MacFarlane, ``New Thermal Neutron Scattering Files
  for ENDF/B-VI Release 2,'' Los Alamos National Laboratory report
  LA-12639-MS (1994).

\bibitem{ICANS} R. E. MacFarlane, ''Cold Moderator Scattering
  Kernels,'' in {\it Advanced Neutron Sources 1988, Proceedings of
  the 10th Meeting of the International Collaboration on Advanced
  Neutron Sources (ICANS-X) held at Los Alamos, 3--7 October 1988},
  Institute of Physics Conferences Series Number 97 (Institute of
  Physics, Bristol and New York, 1988), p.157.

\bibitem{coldworkshop} R. E. MacFarlane, ``Cold Moderator Scattering
  Kernels,'' presented at the International Workshop on Cold Neutron
  Sources, March 5--8, 1990, Los Alamos, New Mexico, Los Alamos
  National Laboratory report LA-12146-C (August 1991).

\bibitem{Parks} D. E. Parks,
  M.  S. Nelkin, J. R. Beyster, and N. F. Wikner,
  {\it Slow Neutron Scattering and Thermalization}, W. A. Benjamin, Inc.
  (New York, 1970).

\bibitem{Wycoff} R. W. G. Wycoff, Crystal Structures, John Wiley \& Sons,
  New York, London, 1963.

\bibitem{YK} J. A. Young and J. U. Koppel, Phys. Rev. \underline{135},
  A603(1964).

\bibitem{EHW} P. A. Egelstaff, B. C. Haywood, and F. J. Webb,
  Proc. Phys. Soc., \underline{90}, 681(1967).

\bibitem{Schott} Wolfgang Schott, Z. Physik \underline{231}, 243(1970).

\bibitem{Utsuro} Masahiko Utsuro, Z. Physik B \underline{27}, 111(1977).

\bibitem{Keinert} J. Keinert and J. Sax, ``Investigation of Neutron Scattering
  Dynamics in Liquid Hydrogen and Deuterium for Cold Neutron Sources,''
  Kerntechnik \underline{51}, 19 (1987).

\bibitem{mattes} M. Mattes and J. Keinert, ``Thermal Neutron Scattering
  Data for the Moderator Materials H2O, D2O and ZrHx in ENDF-6 Format
  and as ACE Library for MCNP(X)", IAEA INDC(NDS)-0470, April 2005.

\bibitem{Dritsa} M. Dritsa, report from misc. OECD countries to EANDC,
  No. 63, 1967.

\bibitem{Heinloth} K. Heinloth, Zeitscrift fur Physik {\bf 163},
  218 (1961).

\bibitem{SmithRR} R. R. Smith, private communication held at BNL, 1953.

\bibitem{Walton} W. Walton, private communication held at BNL, 1960.

\bibitem{Springer} V. T. Springer, Nucleonik {\bf 3}, 110 (1961).

\bibitem{Simpson} O. D. Simpson, J. Nuclear Instruments and Methods
  {\bf 30}, 293 (1964).

\bibitem{Russell} J. L. Russell, Jr., General Atomic report 7581, 1966.

\bibitem{HB} Y. D. Harker and R. M. Brugger,
  ``Investigation of the Low-Temperature Phase Transitions in Solid
  Methane by Slow Neutron Inelastic Scattering,'' J. Chem. Phys.
  \underline{46}, 2201(1967).

\bibitem{AYY} A. K. Agrawal and S. Yip, ``Slow-Neutron
  Scattering by Molecular Liquids,'' Nucl. Sci. Eng. \underline{37},
  368(1969).

\bibitem{Gordon} R. G. Gordon, J. Chem. Phys. \underline{43}, 1307(1965).

\bibitem{Ewing} G. E. Ewing, J. Chem. Phys. \underline{40}, 179(1964).

\bibitem{data} These data were obtained from the paper of Keinert
  and Sax\cite{Keinert} and converted to the numbers shown by
  subtracting the hydrogen absorption cross section.  The ortho
  cross sections were converted from normal $^{2}$H cross sections
  using the theoretical ortho and para cross sections to produce
  a conversion factor.  The original references are G. L. Squires
  and A. T. Stewart, Proc. Roy. Soc. \underline{A230}, 19(1955),
  W. L. Whittemore and A. W. McReynolds, ``Differential Neutron
  Thermalization,'' General Atomic report GA-2505 (1961), and
  W. D. Seiffert, ``Messung der Streuquerschnitte von fl\"{u}ssigem
  und festem Wasserstoff, Deuterium, und Deuteriumhydrid f\"{u}r
  thermische Neutronen,'' EUR-5566d (1970).

\bibitem{recipes} W. H. Press, B. P. Flannery, Saul A. Teukolsky,
  and William T. Vetterling, {\it Numerical Recipes in C}, Cambridge
  University Press, 1988.

\bibitem{UPDATE} ``UPDATE, VERSION 1, Reference Manual,''
  Revision 11/23/81, Control Data Corporation report 60449900
  (November 1981).

\bibitem{ICSBEP} J. B. Briggs, {\it et al}, ``International Handbook
  of Evaluated Criticality Safety Benchmark Experiments,'' Tech. Rep.
  NEA/NSC/DOC(95)04/I, Nuclear Energy Agency, Paris, France, 2004.  Note
  that this Handbook is updated annually.

\bibitem{PREPRO} D. E. Cullen, ``PREPRO 2007: 2007 ENDF/B Pre-Processing
  Codes,'' IAEA-NDS-39, Rev. 13, Nuclear Data Section, International
  Atomic Energy Agency, Vienna, Austria.

\bibitem{sublet} J. C. Sublet, R. N. Blomquest,
  S. Goluoglu, and R. E. MacFarlane, ``Unresolved Resonance
  Range Cross Section Probabiliy and Self Shielding Factors,''
  CEA-R-6227 (July 2009).

\bibitem{fast} D. E. Cullen, P. Brown, E. Lent,
  R. MacFarlane, and S. McKinley, ``Criticality Calculations
  Using LANL and LLNL Neutron Transport Codes,'' Lawrence Livermore
  National Laboratory report UCRL-TR-237333 (November 22, 2007).

\bibitem{big10} R. E. MacFarlane, R. M. Blomquist (ANL),
  D. E. Cullen (LLNL), and J. C. Sublet (CEA),
  ``A Code Comparison Study for the Bigten Critical Assembly,''
  Los Alamos National Laboratory report LA-UR-08-4668 (August 2008).

\bibitem{GIT} Scott Chacon and Ben Straub, {\it Pro Git}, Apress,
  Berkeley, CA, USA, 2nd edition.  Downloadable e-book versions
  are available at \url{https://git-scm.com/book/en/v2}.

\end{singlespace}
\end{thebibliography}

\cleardoublepage

\printindex

\end{document}

